Alleviating Topology Mismatch in Distributed Overlay Networks: A Survey

### 2014-01-27 intro.tex
2012.10.20 (Vassilis) Images reworked. Text alligned with the new images.
2012.10.24 (Vassilis) Changed paper title. Text reworked (content clarifications added, typos corrected, broken references fixed).

Two ``killer'' applications of the peer-to-peer paradigm are, but not limited
to, file-sharing and telephony. P2P file-sharing was originally introduced by
Napster (see Section~\ref{section:background}), which by 2001 was widely
acknowledged as the ``fastest growing Internet application ever" with 26 million
users sharing 80 million songs at the height of its popularity. Other similar
applications quickly followed suit including Gnutella and its descendent,
Limewire, enjoying 3 million concurrently connected peers by the end of 2006
and BitTorrent, connecting over 150 million users. P2P telephony has seen
explosive growth with the advent of Skype. Introduced in 2003, Skype has
experienced explosive popularity with more that 650 million users in 2011
\cite{skypetotalusers} and 35 million skypers concurrently online on the 5th of
March 2012 \cite{skypesymusers}.

@@

Peer-to-peer systems are implemented using overlay networks. An \emph{overlay
network} is a virtual system of nodes and logical links, that is built on top of
an existing network with the aim of providing a service (such as content
distribution) that is not available in the existing one. Nodes in the overlay,
communicate via virtual connections, each of which may correspond to a path of
possibly many physical links in the underlying network.
Figure~\ref{figure:overlay} illustrates a simple four-node overlay constructed
over a wide-area network.

\begin{figure}
\centering
  \includegraphics[scale=0.5]{img/p2p.pdf}
\caption{An example overlay network.}
\label{figure:overlay}
\end{figure}

@@

If the overlay network is constructed such that the topology does not match the 
underlying IP topology well, the resulting \emph{topology mismatch} (See
Section~\ref{section:background} for a more formal definition) can result in two
significant problems. First, application performance (e.g. search query latency
or jitter) can be adversely affected since application traffic must flow over a
larger number of IP hops than if the overlay were constructed more wisely.
Second, studies have shown that P2P applications contribute the largest portion
of overall Internet traffic
\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with
some reporting that more than 60\% of it, is
P2P-related~\cite{cachelogic,ipoque2007,ipoque2009} and others claiming that it
is expected to reach almost 8 petabytes per month by the end of
2012~\cite{multinteligence}. This is a significant burden for Internet Service
Providers (ISPs) who must handle all this traffic to destination nodes sitting
at the edge of the Internet. If the overlay topology of a peer-to-peer system is
not efficintly constructed, then this burden to the Internet's backbone
infrastructure could be even greater. For example, traffic might have to flow
back and forth several times between two neighboring ISPs as it travels from its
source node in the overlay to its destination node. Thus, it is fundamentally
important that P2P overlay networks are constructed and maintained so that
their topology matches as closely as possible the underlying IP topology.

For over a decade, several tens of researchers have published profusely on the 
topology mismatch problem.  In this paper, we survey this research area and
provide a taxonomy of the proposed solutions. We point out synergies as well as
similarities and differences in the published approaches. Ultimately, our goal
is to help readers sift through the vast amount of literature, to help them
understand the advantages and disadvantages of each of each work, and to provide
them with enough perspective so that when the need arises, they are able to
select amongst the different approaches for their particular application.

The rest of the paper is as follows: In Section~\ref{section:background}, we
provide some background information on peer-to-peer overlay architectures
including centralized, decentralized unstructured and decentralized structured
peer-to-peer systems. We also formally define the problem of topology mismatch
between an overlay and its underlying physical network and elaborate on the
rationale behind our coarse grained evaluation of the techinques presented
throughout this survey. In Sections~\ref{section:unstructured} and
\ref{section:structured}, we present the various research efforts for
decentralized unstructured and decentralized structured peer-to-peer systems,
respectively. Finally, with Section~\ref{section:conclusion}, we
conclude.

>>>>

Two ``killer'' applications of the peer-to-peer paradigm are, but not limited
to, file-sharing and telephony. P2P file-sharing was originally introduced by
Napster (see Section~\ref{section:background}), which by 2001 was widely
acknowledged as the ``fastest growing Internet application ever" with 26 million
users sharing 80 million songs at the height of its popularity. Other similar
applications quickly followed suit including Gnutella and its descendant,
Limewire, enjoying 3 million concurrently connected peers by the end of 2006
and BitTorrent, connecting over 150 million users. P2P telephony has seen
explosive growth with the advent of Skype. Since its introduction in 2003,
Skype has become extremely popular with more than 650 million users in 2011
\cite{skypetotalusers} and 35 million skypers concurrently online on the 5th of
March 2012 \cite{skypesymusers}.

@@

Peer-to-peer systems are implemented using overlay networks. An \emph{overlay
network} is a virtual system of nodes and logical links, that is built on top of
an existing network intending to provide added value to it (e.g., offering QoS
content distribution over the best effort Internet network). Nodes in the
overlay communicate via virtual connections each of which may correspond to a
path of possibly many physical links in the underlying network.
Figure~\ref{figure:overlay} illustrates a simple four-node overlay constructed
over a wide-area network.

\begin{figure}
\centering
  \includegraphics[scale=0.7]{img/pdf/under-over-lay.pdf}
\caption{An example overlay network.}
\label{figure:overlay}
\end{figure}

@@

If the overlay network is constructed such that it does not match the underlying
topology well, the resulting \emph{topology mismatch} (See
Section~\ref{section:background} for a more formal definition) can result in two
significant problems. First, the performance of the application per se, can be
adversely affected since application traffic must flow over a larger,
redundunt, number of physical hops resulting in poor user experience like
extensive search latency or jitter. Second, can cumulatively degrade the
performance of the whole underlying infrastructure, effecting other applications
as well. For example, studies have shown that the extreme popularity of P2P
applications amply deployed everywhere nowadays, result in them contributing the
largest portion of overall Internet traffic
\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with
some reporting that more than 60\% of it, is
P2P-related~\cite{cachelogic,ipoque2007,ipoque2009} and others claiming that it
is expected to reach almost 8 petabytes per month by the end of
2012~\cite{multinteligence}. This is a significant burden for Internet Service
Providers (ISPs) who must handle all this traffic to destination nodes sitting
at the edge of the Internet. If the overlay topology of a peer-to-peer system is
not efficiently constructed, then this burden to the Internet's backbone
infrastructure could be even greater. For example, traffic might have to flow
back and forth several times between two neighboring ISPs as it travels from its
source node in the overlay to its destination node. Thus, it is fundamentally
important that P2P overlay networks are constructed and maintained so that
their topology matches as closely as possible the underlying IP topology.

For over a decade, several researchers have published profusely on the topology
mismatch problem.  In this paper, we survey this research area and provide a
taxonomy of the proposed solutions. We point out synergies as well as
similarities and differences in the published approaches. Ultimately, our goal
is to help readers sift through the vast amount of literature, to help them
understand the advantages and disadvantages of each of each work, and to provide
them with enough perspective so that when the need arises, they are able to
select amongst the different approaches for their particular application.

The rest of the paper is organized as follows: In
Section~\ref{section:background}, we provide background information on
peer-to-peer overlay architectures including centralized, decentralized
unstructured and decentralized structured peer-to-peer systems. We also formally
define the problem of topology mismatch between an overlay and its underlying
physical network and elaborate on the rationale behind our coarse grained
evaluation of the techniques presented throughout this survey. In
Sections~\ref{section:unstructured} and \ref{section:structured}, we present the
various research efforts for decentralized unstructured and decentralized
structured peer-to-peer systems, respectively. Finally, with
Section~\ref{section:conclusion}, we conclude the paper.

### 2014-01-27 mard-survey.tex
2012.10.20 (Vassilis) Images reworked. Text alligned with the new images.
2012.10.24 (Vassilis) Changed paper title. Text reworked (content clarifications added, typos corrected, broken references fixed).

% packages
\usepackage[utf-8]{inputenc}
\usepackage{paralist}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{colortbl}
\usepackage{color,soul}
\usepackage{longtable}
\usepackage{wasysym}
\usepackage{pdflscape}
%\usepackage{tikz}
\usepackage{pgfplots}

@@

\title{Alleviating the Topology Mismatch Problem in Unstructured and Structured
Overlay Networks: A Survey}
% TODO: We added here "P2P". I'm not sure, because for example Narada is not P2P
% TODO: I deleted "P2P"

@@

\begin{abstract}
Peer-to-peer (P2P) systems is a rapidly growing application architecture. The
overlay network abstraction on top of a best-effort infrastructure provides a
powerful tool to realize a plethora of ever-wanted exotic features like
anonymity, high availability and robustness, load balancing, quality-of-service
(QoS) at the application layer and many more. Unfortunately weaknesses of early
effectuations constrained applications from unleashing the full potential of the
paradigm. One of them is the topology mismatch problem between the overlay and
the physical underlying network, a problem that imposes a huge amount of
unnecessary stress to the network resources and, thus, impede applications from
scaling. In this survey we investigate on the recent research developments
towards the alleviation of this problem, in both unstructured and structured
overlay architectures. We visit the various algorithms available in the
lietrature, focusing on the presentation of how each of them tries to tackle the
problem. Additionally, we provide a finer grained categorisation of the
approaches based on the generic type of methodology they use in order to
accomplish this goal.
% TODO: Review this after the comparison sections are fixed.
Finally we discuss the advantages, disadvantages and novelties introduced by the
algotithms and offer pictorial comparisons for them in order to further
strengthen and organise understanding of the information and knowledge cited, in
an attempt to make this survey a one-stop-shop reference for anyone interested
in the topic, from researchers and students, to system architects and
developers.
\end{abstract}

@@

% \item Proximity Neighbour Selection\\
%Finally, the third approach, constructs the routing tables using proximity
%knowledge. Tapestry and Pastry's mechanisms of routing table
%maintenance try to minimize the distance to nodes appearing in a peer's routing
%table. Since routing is based on longest node ID prefix match, messages 
%are gradually forwarded to nearby nodes at each routing step.
%\cite{castro_proximityp2p_2002} argues on how Pastry exploits proximity
%neighbour selection in order to create a scheme that is (more) location-aware
%compared to the other well-known DHTs (CAN, Chord).
%
%\end{itemize}
%
%{\sethlcolor{yellow}\hl{
%HA: Possible criteria: based on which protocol (eg Gnutella, CAN etc), peer
%selection (topology selection, cluster, cache etc), supports dynamic update,
%runtimes}}
%%\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Forwarding} &
%\textbf{Cache} & \textbf{Overlay optimization} & \textbf{Proximity information}
%& \textbf{Base protocol} & \textbf{Dynamic update} & \textbf{Runtime} \\
%
%
%{\sethlcolor{yellow}\hl{
%HA: Possible criteria2: Overlay optimization structure, base protocol (eg
%Gnutella, CAN etc), dynamic update, runtimes, scalability}}
%
%{\sethlcolor{yellow}\hl{
%HA: maybe add also the year of publication, and see if there is a pattern in
%terms of the method and the year??}}

>>>>

% packages
\usepackage[utf-8]{inputenc}
\usepackage{paralist}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{colortbl}
\usepackage{color,soul}
\usepackage{longtable}
\usepackage{wasysym}
\usepackage{pdflscape}
%\usepackage{tikz}
\usepackage{pgfplots}
\usetikzlibrary{patterns}

@@

\title{Alleviating Topology Mismatch in Distributed Overlay Networks: A Survey}
% TODO: We added here "P2P". I'm not sure, because for example Narada is not P2P
% TODO: I deleted "P2P"

@@

\begin{abstract}
Peer-to-peer (P2P) systems is a rapidly growing application architecture. The
overlay network abstraction on top of a best-effort infrastructure provides a
powerful tool to realize a plethora of ever-wanted exotic features like
anonymity, high availability and robustness, load balancing, quality-of-service
(QoS) at the application layer and many more. Unfortunately weaknesses of early
effectuations constrained applications from unleashing the full potential of the
paradigm. One of them is the topology mismatch problem between the overlay and
the physical underlying network, a problem that imposes a huge amount of
unnecessary stress to the network resources and, thus, impede applications from
scaling. In this survey we investigate on the recent research developments
towards the alleviation of this problem, in both unstructured and structured
overlay architectures. We visit the various algorithms available in the
literature, focusing on the presentation of how each of them tries to tackle the
problem. Additionally, we provide a finer grained categorization of the
approaches based on the generic type of methodology they use in order to
accomplish this goal.
% TODO: Review this after the comparison sections are fixed.
Finally we discuss the advantages, disadvantages and novelties introduced by the
algorithms and offer pictorial comparisons for them in order to further
strengthen and organize understanding of the information and knowledge cited, in
an attempt to make this survey a one-stop-shop reference for anyone interested
in the topic, from researchers and students, to system architects and
developers.
\end{abstract}

@@

% \item Proximity Neighbor Selection\\
%Finally, the third approach, constructs the routing tables using proximity
%knowledge. Tapestry and Pastry's mechanisms of routing table
%maintenance try to minimize the distance to nodes appearing in a peer's routing
%table. Since routing is based on longest node ID prefix match, messages 
%are gradually forwarded to nearby nodes at each routing step.
%\cite{castro_proximityp2p_2002} argues on how Pastry exploits proximity
%neighbor selection in order to create a scheme that is (more) location-aware
%compared to the other well-known DHTs (CAN, Chord).
%
%\end{itemize}
%
%{\sethlcolor{yellow}\hl{
%HA: Possible criteria: based on which protocol (eg Gnutella, CAN etc), peer
%selection (topology selection, cluster, cache etc), supports dynamic update,
%runtimes}}
%%\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Forwarding} &
%\textbf{Cache} & \textbf{Overlay optimization} & \textbf{Proximity information}
%& \textbf{Base protocol} & \textbf{Dynamic update} & \textbf{Runtime} \\
%
%
%{\sethlcolor{yellow}\hl{
%HA: Possible criteria2: Overlay optimization structure, base protocol (eg
%Gnutella, CAN etc), dynamic update, runtimes, scalability}}
%
%{\sethlcolor{yellow}\hl{
%HA: maybe add also the year of publication, and see if there is a pattern in
%terms of the method and the year??}}
